\section{Model Architectures}
\label{sec:methodology}

We study two families of detectors, chosen to represent complementary
points on the accuracy--efficiency Pareto frontier.

\subsection{PicoDet-S}

PicoDet~\citep{yu2021picodet} is an anchor-free single-stage detector
designed for real-time mobile inference.  The PicoDet-S variant
(``small'') uses:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Backbone:} LCNet-0.75$\times$~\citep{cui2021pp},
    a lightweight depthwise-separable network with squeeze-excitation
    blocks, pre-trained on ImageNet.
  \item \textbf{Neck:} LCPAN, a lightweight path-aggregation network
    combining top-down and bottom-up feature fusion.
  \item \textbf{Head:} TOOD-style task-aligned classification and
    regression heads with Varifocal Loss (VFL) and Distribution Focal
    Loss (DFL).
  \item \textbf{Input resolution:} 320$\times$320.
  \item \textbf{Parameters:} $\approx$\,1.2\,M; \textbf{model size:}
    4.4\,MB (FP32 ONNX).
\end{itemize}

The anchor-free design is particularly advantageous on small datasets
because it eliminates the need for hand-tuned anchor configurations
that must match the training data distribution.
COCO pre-training weights are used for the backbone and neck; the
classification head is re-initialised for two output classes.

\subsection{YOLOv3-MobileNetV1}

YOLOv3~\citep{redmon2018yolov3} is an anchor-based multi-scale detector.
We use the official PaddleDetection implementation with
MobileNetV1~\citep{howard2017mobilenets} as the backbone:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Backbone:} MobileNetV1 pre-trained on COCO (80-class head
    discarded; feature extractor retained).
  \item \textbf{Detection head:} Three-scale prediction ($52\!\times\!52$,
    $26\!\times\!26$, $13\!\times\!13$), nine default anchors
    (COCO-derived).
  \item \textbf{Input resolution:} 608$\times$608.
  \item \textbf{Parameters:} $\approx$\,23\,M; \textbf{model size:}
    92\,MB (FP32 ONNX).
  \item \textbf{Post-processing:} \texttt{multiclass\_nms3}
    (PaddleDetection custom operator).
\end{itemize}

\subsection{Architecture Comparison}

\begin{table}[H]
\centering
\caption{Architecture comparison between the two detector families.}
\label{tab:arch}
\begin{tabularx}{\textwidth}{lXX}
\toprule
\textbf{Property} & \textbf{PicoDet-S} & \textbf{YOLOv3-MobileNetV1} \\
\midrule
Detector type       & Anchor-free      & Anchor-based \\
Backbone            & LCNet-0.75$\times$ & MobileNetV1 \\
Input size          & 320$\times$320   & 608$\times$608 \\
Parameters          & $\approx$1.2\,M  & $\approx$23\,M \\
ONNX size (FP32)    & 4.4\,MB          & 92\,MB \\
FPS on Tesla T4     & $\sim$78         & $\sim$35 \\
CoreML compatibility & \checkmark       & $\times$ (custom NMS op) \\
COCO pre-train      & LCNet backbone + LCPAN neck & MobileNetV1 backbone \\
\bottomrule
\end{tabularx}
\end{table}
